---
title: "Answer to Reproducble Research: Peer Assessment 1"
author: "JC"
date: "12/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1

R Code for Reading the dataset and/or processing the data. First 'activity.zip' file is unziped with command unzip. File activity.csv is read using read.csv command. Summary and Str shows that we have NAs in variable step.

```{r}
unzip("./activity.zip")
DataActivity <- read.csv("./activity.csv")
summary(DataActivity)
str(DataActivity)
```

## Part 2

Histogram of the total number of steps taken each day. Aggregate creates a dataframe using steps into date subsets.It has ignored the NAs in steps. 

```{r}
StepsDay <- aggregate(steps ~ date, DataActivity, FUN = sum, na.rm = TRUE)
hist(StepsDay$steps)
```

## Part 3
Mean and median number of steps taken each day using the frequencies. 

```{r}
meanStepsPerDay <- mean(StepsDay$steps)
meanStepsPerDay # Mean
medianStepsPerDay <- median(StepsDay$steps)
medianStepsPerDay # Median
```
## Part 4

Time series plot of the average number of steps taken. 

```{r}
StepsPerInt<-aggregate(steps~interval, data=DataActivity, mean, na.rm=TRUE)
plot(steps~interval, data=StepsPerInt, type="l")
```

## Part 5

The 5-minute interval that, on average, contains the maximum number of steps. Which.max calculates the index in the dataframe that has the maximum number of steps (104).

```{r}
MaxInt <- StepsPerInt[which.max(StepsPerInt$steps),]$interval
MaxInt
```
## Part 6

Code to describe and show a strategy for imputing missing data using the mean per interval.

```{r}
MeanPerInt <- function(interval) {
  StepsPerInt[StepsPerInt$interval==interval,]$steps
}

NewDataNoNA <- DataActivity
for(i in 1:nrow(NewDataNoNA)){
  if(is.na(NewDataNoNA[i,]$steps)){
    NewDataNoNA[i,]$steps <- MeanPerInt(NewDataNoNA[i,]$interval)
  }
}
```
## Part 7

Histogram of the total number of steps taken each day after missing values are imputed. 

```{r}
NewDataNoNAHist <- aggregate(steps ~ date, NewDataNoNA, sum)
hist(NewDataNoNAHist$steps)
```

## Part 8

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends. 

```{r}
NewDataNoNA$date <- as.Date(strptime(NewDataNoNA$date, format="%Y-%m-%d"))
NewDataNoNA$day <- weekdays(NewDataNoNA$date)
for (i in 1:nrow(NewDataNoNA)) {
    if (NewDataNoNA[i,]$day %in% c("Saturday","Sunday")) {
        NewDataNoNA[i,]$day<-"weekend"
    }
    else{
        NewDataNoNA[i,]$day<-"weekday"
    }
}
StepsByDay <- aggregate(NewDataNoNA$steps ~ NewDataNoNA$interval + NewDataNoNA$day, NewDataNoNA, mean)
head(StepsByDay)

names(StepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, StepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```


